---
title: "R Notebook"
output:
  html_document: default
  html_notebook: default
---
# 0 count trials were replaced with the average count of 0.5 [based on limit of detection/2], as 0 cannot be log transferred for statistical analysis.

# 1. load library
```{r}
library(nlme)
library(car)
library(multcompView)
library(multcomp)
library(lsmeans)
library(rcompanion)
library(dplyr)
library(reshape2)
library(ggplot2)
library(FSA)
library(ggpattern)
```


# 2. Milk (Repeated Measures ANOVA)
## whole milk
```{r}
options(width=120)
#analysis ref:http://rcompanion.org/handbook/I_09.html

wm<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/wholemilk.csv",header = T)
wm1<-subset(wm, Time!=0)

#The autocorrelation structure
model.a = gls(log10 ~ concen + Time + concen*Time,
             data=wm)
ACF(model.a,
    form = ~ Time | rep)


# model
model = gls(log10 ~ concen + Time + concen*Time,
            correlation = corAR1(form = ~ Time | rep,
                                 value = 0.1988),
            data=wm,
            method="REML")

summary(model)

# variance analysis
Anova(model)

# residule 
x = residuals(model)
library(rcompanion)
plotNormalHistogram(x)

#posthoc
marginal = lsmeans(model,~concen*Time, cov.reduce = FALSE)

pairs = as.data.frame(pairs(marginal,
      adjust="tukey"))
pairs

Sum = Summarize(log10 ~ concen*Time ,
                    data   = wm1,
                    conf   = 0.95,
                    digits = 3,
                    traditional = FALSE,
                    percentile  = TRUE)

Sum

# add asterisk compared to control
sig=c(" "," "," "," ",
      " "," ","*"," ",
      " "," ","*"," ",
      " "," ","***"," ",
      " "," "," "," ",
      " "," "," "," ")

sum1<-cbind(Sum,sig)

sum1$concen.ord<-factor(sum1$concen,c("control","1000","1400","2000"))

p=ggplot(sum1, aes(x = Time,y = mean, shape = concen.ord)) + 
    geom_point(size=3) +  
    scale_shape_manual(values=c(15,16,17,0),labels = c("SDW control","1000 mg/L","1400 mg/L","2000  mg/L")) + 
    geom_line() +
    geom_errorbar(aes(ymin=mean-sd,ymax=mean+sd),width=.3, size=0.7) +
    geom_text(aes(label = sum1$sig), vjust=2) + 
    theme_bw() +
    geom_hline(yintercept=4, linetype="dashed") +
    theme(legend.title=element_blank(),panel.grid=element_blank()) +
    ylab(expression(~italic("L. monocytogenes")~" counts (log CFU/mL)")) + xlab("Time (d)") + ylim(0,8.5)

p

#old figure setting
#p=ggplot(sum1, aes(x = Time,y = mean, shape = concen.ord, color=concen.ord)) + 
#    geom_point(size=4) + geom_line() +
#    geom_errorbar(aes(ymin=mean-sd,ymax=mean+sd),
#                      width=.3, size=0.7)  + scale_shape_discrete(labels = c("Control","1000 #mg/L","1400 mg/L","2000 mg/L")) + scale_color_grey(labels = c("Control","1000 mg/L","1400 #mg/L","2000 mg/L"),start = 0, end = 0.7) +
#    geom_text(aes(label = sum1$sig), hjust=-0.4) +
#    theme_bw() +
#    theme(legend.title=element_blank(),panel.grid=element_blank()) +
#    ylab(expression(~italic("L. monocytogenes")~" counts (log CFU/mL)")) + xlab("Time (d)") + #ylim(0,8.5)

#ggsave(p,file="~/Desktop/lm.wm.jpeg",dpi=300,height = 4, width = 6)
```
## skim milk
```{r}
sm<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/skimmilk.csv",header = T)
sm1<-subset(sm, Time!=0)

#The autocorrelation structure
model.a = gls(log10 ~ concen + Time + concen*Time,
             data=sm1)
ACF(model.a,
    form = ~ Time | rep)


# model
model = gls(log10 ~ concen + Time + concen*Time,
            correlation = corAR1(form = ~ Time | rep,
                                 value = 0.1753),
            data=sm,
            method="REML")

summary(model)

# variance analysis
Anova(model)

# residule 
x = residuals(model)
plotNormalHistogram(x)

#posthoc
marginal = lsmeans(model,~concen*Time, cov.reduce = FALSE)
pairs = as.data.frame(pairs(marginal,
      adjust="tukey"))
pairs

#summary
Sum = Summarize(log10 ~ concen*Time ,
                    data   = sm1,
                    conf   = 0.95,
                    digits = 3,
                    traditional = FALSE,
                    percentile  = TRUE)

Sum

# add asterisk compared to control
sig=c("**      ","**      ","***      "," "," "," ",
      "***      ","***      ","***      "," "," "," ",
      "***","***","***"," "," "," ",
      "***","***      ","***      "," "," "," ",
      " ","***","***"," "," "," ",
      " "," "," "," "," "," ")

sum1<-cbind(Sum,sig)

sum1$concen.ord<-factor(sum1$concen,c("control","500","700","1000","1300","1500"))

rm.bl.limit<-sum1
rm.bl.limit$mean[rm.bl.limit$mean == "-0.3"] <- "NA"
rm.bl.limit$mean<-as.numeric(rm.bl.limit$mean)

p=ggplot(rm.bl.limit, aes(x = Time,y = mean, shape = concen.ord)) + 
    geom_point(size=3) +  
    scale_shape_manual(values=c(15,16,17,0,1,2),labels = c("SDW control","500 mg/L","700 mg/L", "1000 mg/L","1300 mg/L","1500 mg/L")) + 
    geom_line() +
    geom_errorbar(aes(ymin=mean-sd,ymax=mean+sd),width=.3, size=0.7) +
    geom_text(aes(label = rm.bl.limit$sig), vjust=2.5) + 
    geom_text(aes(x=21, y=0.2,label = "^"), size=5, vjust=1.5) + 
    theme_bw() +
    geom_hline(yintercept=4, linetype="dashed") +
    theme(legend.title=element_blank(),panel.grid=element_blank()) +
    ylab(expression(~italic("L. monocytogenes")~" counts (log CFU/mL)")) + xlab("Time (d)") + ylim(-0.2,8.5)

p

#p=ggplot(sum1, aes(x = Time,y = mean, shape = concen.ord, color=concen.ord)) + 
#    geom_point(size=4) + geom_line() +
#    geom_errorbar(aes(ymin=mean-sd,ymax=mean+sd),
#                      width=.3, size=0.7)  +      
#    scale_color_grey(labels = c("Control mg/L","500 mg/L","700 mg/L", "1000 mg/L","1300 #mg/L","1500 mg/L")) + 
#    scale_shape_discrete(labels = c("Control mg/L","500 mg/L","700 mg/L", "1000 mg/L","1300 #mg/L","1500 mg/L")) +
#    geom_text(aes(label = sum1$sig), hjust=-0.4) +
#    theme_bw() +
#    theme(legend.title=element_blank(),panel.grid=element_blank()) +
#    ylab(expression(~italic("L. monocytogenes")~" counts (log CFU/mL)")) + xlab("Time (d)") + #ylim(-1,8.5) + geom_hline(yintercept=0, linetype="dashed", color = "grey", size=1)

#ggsave(p,file="~/Desktop/lm.sm1.jpeg",dpi=300,height = 4, width = 6)
```

```{r,include=FALSE}
skimmilk<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/skimmilk.csv",header = T)

skimmilk2<-subset(skimmilk,!(skimmilk$concen=="700"|skimmilk$concen=="1300"))

sm=subset(skimmilk2,skimmilk2$Time!="0")
sm$Time = factor(sm$Time,
                   levels=c("1", "3", "5", "7","14","21"))

###### skim milk only keep 500, 1000, and 1500. 700 and 1300 not shown
# model
model = lm(log10 ~ concen*Time + rep,
           data = sm)

# variance analysis
library(car)
Anova(model,
      type = "II")

# residule 
x = residuals(model)

library(rcompanion)

plotNormalHistogram(x)

# post hoc
library(multcompView)
library(lsmeans)

marginal = lsmeans(model,
                   ~ concen*Time)

pairs = as.data.frame(pairs(marginal,
      adjust="tukey"))
pairs

CLD = cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey") 
CLD

library(rcompanion)
library(FSA)

Sum = Summarize(log10 ~ concen + Time,
          data=skimmilk2,
          digits=3)

Sum

# add asterisk compared to control
sig=c(" "," "," "," ",
      "***","***"," "," ",
      "***","***","***"," ",
      "***","***","***"," ",
      "***","***"," "," ",
      "***","***"," "," ",
      " ","***"," "," ")

sum1<-cbind(Sum,sig)

sum1$concen.ord<-factor(sum1$concen,c("500","1000","1500","control"))

pd = position_dodge(.2)

p=ggplot(sum1, aes(x = Time,y = mean, color = concen.ord)) + geom_point(shape=15, size=4, position=pd) + geom_line() +
    geom_errorbar(aes(ymin=mean-sd,
                      ymax=mean+sd),
                   width=.2, size=0.7, position=pd) + scale_color_grey(labels = c("500 mg/L",  "1000 mg/L","1500 mg/L","Control")) +
    geom_text(aes(label = sum1$sig,y=mean), hjust=-0.5) +
    theme_bw() +
    theme(legend.title=element_blank(),panel.grid=element_blank()) +
    ylab(expression(~italic("L. monocytogenes")~" counts (log CFU/mL)")) + xlab("Time (d)") + ylim(-1,8.5) + geom_hline(yintercept=0, linetype="dashed", color = "grey", size=1)

p

#ggsave(p,file="~/Desktop/lm.sm2.png",height = 7, width = 10)

```




# 3. Cheese (Two-way ANOVA, followed by one-way ANOVA)
## cheese
```{r}
# analysis ref: https://rcompanion.org/handbook/I_08.html

cheese1<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/cheese.csv",header = T)

# combine technical reps and get the average
cheese<-aggregate(cheese1[, 6], list(cheese1$reps, cheese1$sample,cheese1$day,cheese1$treatment, cheese1$rep),mean)
colnames(cheese) <- c("reps","sample","day","treatment","rep","log10")

#remove day 28 and 35 for analysis, cause we are not gonna report them
cheese <- subset(cheese, cheese$day!="28")
cheese <- subset(cheese, cheese$day!="35")

# model
model = lm(log10 ~ day+treatment+day*treatment, data=cheese)
# two way anova
Anova(model, type = "II")
# residule 
x = residuals(model)
plotNormalHistogram(x)
#posthoc
marginal = lsmeans(model,
                   ~ treatment)
pairs(marginal,
      adjust="tukey")

#one way anova at each time point
# day1
aov=aov(log10 ~ treatment, data=subset(cheese,cheese$day=="1"))
marginal = lsmeans(aov,~ treatment)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
# day7
aov=aov(log10 ~ treatment, data=subset(cheese,cheese$day=="7"))
marginal = lsmeans(aov,~ treatment)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
# day14
aov=aov(log10 ~ treatment, data=subset(cheese,cheese$day=="14"))
marginal = lsmeans(aov,~ treatment)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
# day21
aov=aov(log10 ~ treatment, data=subset(cheese,cheese$day=="21"))
marginal = lsmeans(aov,~ treatment)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
```
## cheese graph
```{r}
cheese.pic<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/figure/cheese.csv",header = T)

cheese.pic$concen.ord<-factor(cheese.pic$concen,c("control","DI","5","10","15","20"))

#p<-ggplot(cheese.pic, aes(factor(day), y=ave,fill=concen.ord)) +
#  geom_bar(stat="identity", position = "dodge",color="black") +
#  geom_errorbar(aes(ymin = ave-std, ymax = ave+std), width =0.25,position=position_dodge(0.9)) #+
#  geom_text(aes(label = cheese.pic$group,y=ave+std),vjust=-1,position=position_dodge(0.9))+
#  scale_fill_grey(labels = c("Control","SDW control","5%",  "10%","15%","20%"),start = 0, end #= 1) +
#  ylab(expression(~italic("L. monocytogenes")~"counts (log CFU per gram)")) + xlab("Time (d)") #+
#  theme(legend.title=element_blank(),panel.background = element_blank(), #panel.grid=element_blank(), axis.line=element_line('black'),strip.placement = "outside")

#previous figure was above, but want to add patterns
# install.packages("remotes")
#remotes::install_github("coolbutuseless/ggpattern")

p<-ggplot(cheese.pic, aes(factor(day), y=ave, pattern_type = concen.ord,fill=concen.ord)) +
  geom_bar_pattern(stat="identity", position = "dodge",color="black",pattern = 'magick',pattern_fill="white") +
  scale_pattern_type_manual(values = c('hexagons', 'left45', 'crosshatch', 'right45', 'checkerboard', "left45"),labels = c("Control","SDW control","5%",  "10%","15%","20%")) +
  geom_errorbar(aes(ymin = ave-std, ymax = ave+std), width =0.25,position=position_dodge(0.9)) +
  geom_text(aes(label = cheese.pic$group,y=ave+std),vjust=-1,position=position_dodge(0.9))+
  scale_fill_grey(labels = c("Control","SDW control","5%",  "10%","15%","20%"),start = 0, end = 1) +
  ylab(expression(~italic("L. monocytogenes")~"counts (log CFU per gram)")) + xlab("Time (d)") +
  theme(legend.title=element_blank(),panel.background = element_blank(), panel.grid=element_blank(), axis.line=element_line('black'),strip.placement = "outside")

p
#ggsave(p,file="~/Desktop/lm.cheese.jpeg",height = 5, width = 8)
```


# 4. Biofilm (Two-way ANOVA, followed by one-way ANOVA)
## inhibition
### stainless steel inhibition
```{r}
steel_inhi<- read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/steel_un_inhi.csv",header=T)

s.inhi=subset(steel_inhi,steel_inhi$time!="0")
s.inhi$time = factor(s.inhi$time,
                   levels=c("24","48"))

#linear model
model = lm(log10.per.coupon ~ concentration*time,
           data = s.inhi)

# two way anova
Anova(model,
      type = "II")

# residule normality
x = residuals(model)
plotNormalHistogram(x)

#one way anova at each time point
# 24h
aov=aov(log10.per.coupon ~ concentration, data=subset(s.inhi,s.inhi$time=="24"))
marginal = lsmeans(aov,~ concentration)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
# 48h
aov=aov(log10.per.coupon ~ concentration, data=subset(s.inhi,s.inhi$time=="48"))
marginal = lsmeans(aov,~ concentration)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")

```
#### polystyrene inhibition 
```{r}
poly_inhi<- read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/poly_un_inhi.csv",header=T)

p.inhi=subset(poly_inhi,poly_inhi$time!="0")
p.inhi$time = factor(p.inhi$time,
                   levels=c("24","48"))

#linear model
model = lm(log10_well ~ concentration*time,
           data = p.inhi)

# two way anova
Anova(model,
      type = "II")

# residule normality
x = residuals(model)
plotNormalHistogram(x)

#one way anova at each time point
# 24h
aov=aov(log10_well ~ concentration, data=subset(p.inhi,p.inhi$time=="24"))
marginal = lsmeans(aov,~ concentration)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
# 48h
aov=aov(log10_well ~ concentration, data=subset(p.inhi,p.inhi$time=="48"))
marginal = lsmeans(aov,~ concentration)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
```
### inhi graph
```{r}
inhi.pic<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/3.\ Glycolipid\ project/data/statistics/figure/inhi.csv",header = T)

inhi.pic$concen.ord<-factor(inhi.pic$concen,c("control","1.6"))

p<-ggplot(inhi.pic, aes(factor(time), y=ave,fill=concen.ord)) +
  geom_bar(stat="identity", position = "dodge",color="black")+
  geom_errorbar(aes(ymin = ave-std, ymax = ave+std), width = 0.25,position=position_dodge(0.9))+
  facet_grid(.~material,switch = "both")+
  geom_text(aes(label = inhi.pic$group,y=ave+std),vjust=-1,position=position_dodge(0.9))+
  scale_fill_grey(labels = c("G1 (1.6 mg/L)","SDW control"))+
  ylab(expression(~italic("L. monocytogenes")~"counts (log CFU per well or round)")) + xlab("Time (h)") +
  theme(legend.title=element_blank(),panel.background = element_blank(), panel.grid=element_blank(), axis.line=element_line('black'),strip.placement = "outside")

p
#ggsave(p,file="~/Desktop/lm.inhi.jpeg",dpi = 300, height = 5, width = 6)
```



## inactivation
### stainless steel inactivation
```{r}
steel_inacti<- read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/steel_un_inacti.csv",header=T)

s.inact=subset(steel_inacti,steel_inacti$time!="0")
s.inact$time = factor(s.inact$time,
                   levels=c("1","4"))

#linear model
model = lm(log10per_coupon ~ concentration*time,
           data = s.inact)

# two way anova
Anova(model, type = "II")

# residule normality
x = residuals(model)
plotNormalHistogram(x)

#post hoc
marginal = lsmeans(model,~ concentration)
pairs(marginal,adjust="tukey")

#one way anova at each time point
# 1h
aov=aov(log10per_coupon ~ concentration, data=subset(s.inact,s.inact$time=="1"))
marginal = lsmeans(aov,~ concentration)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")

# 4h
aov=aov(log10per_coupon ~ concentration, data=subset(s.inact,s.inact$time=="4"))
marginal = lsmeans(aov,~ concentration)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
```
### polystyrene inactivation
```{r}
poly_un<- read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/poly_un_inacti.csv",header=T)

p.inact=subset(poly_un,!(poly_un$time_h=="0"|poly_un$concentration_ppm=="22.5"|poly_un$concentration_ppm=="90"))
p.inact$time_h = factor(p.inact$time_h,
                   levels=c("1","4"))

#linear model
model = lm(log10_well ~ concentration_ppm*time_h,
           data = p.inact)

# two way anova
Anova(model, type = "II")

# residule normality
x = residuals(model)
plotNormalHistogram(x)

#post hoc
marginal = lsmeans(model,~ concentration_ppm)
pairs(marginal, adjust="tukey")

#one way anova at each time point
# 1h
aov=aov(log10_well ~ concentration_ppm, data=subset(p.inact,p.inact$time_h=="1"))
marginal = lsmeans(aov,~ concentration_ppm)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")

# 4h
aov=aov(log10_well ~ concentration_ppm, data=subset(p.inact,p.inact$time_h=="4"))
marginal = lsmeans(aov,~ concentration_ppm)
pairs(marginal,adjust="tukey")
cld(marginal,
          alpha   = 0.05,
          Letters = letters,    
          adjust  = "tukey")
```

### inacti graph
```{r}
inacti.pic<-read.csv("/Volumes/efs/CANR/Ansci/D\'Amico\ Lab/Langsun_las17015/Glycolipid\ project/data/statistics/figure/inacti.csv",header = T)

inacti.pic$concen.ord<-factor(inacti.pic$concen,c("PBS","H2O","4.5","45","450"))

p<-ggplot(inacti.pic, aes(factor(time), y=ave,fill=concen.ord,pattern = concen.ord,pattern_angle = concen.ord)) +
  geom_bar_pattern(stat="identity", position = "dodge",color="black",pattern_spacing = 0.025)+
  geom_errorbar(aes(ymin = ave-std, ymax = ave+std), width = 0.25,position=position_dodge(0.9))+
  facet_grid(.~materials,switch = "both")+
  geom_text(aes(label = inacti.pic$group,y=ave+std),vjust=-1,position=position_dodge(0.9))+
  scale_fill_grey(labels = c("PBS control","SDW control","4.5 mg/L","45 mg/L","450 mg/L"),start = 0.3, end = 1)+
  scale_pattern_discrete(labels = c("PBS control","SDW control","4.5 mg/L","45 mg/L","450 mg/L"))+
  scale_pattern_angle_discrete(labels = c("PBS control","SDW control","4.5 mg/L","45 mg/L","450 mg/L"))+
  ylab(expression(~italic("L. monocytogenes")~"counts (log CFU per well or round)")) + xlab("Time (h)") +
  theme(legend.title=element_blank(),panel.background = element_blank(), panel.grid=element_blank(), axis.line=element_line('black'),strip.placement = "outside")+
  ylim(0,8.5)

p

#ggsave(p,file="~/Desktop/lm.inacti.jpeg",dpi=300,height = 5, width = 7)
```

